---
title: "R_packages"
author: "Lauren Trichtinger"
date: "2/2/2022"
output: html_document
---

## Packages in R 

Currently CRAN has 18,858 available packages for R. You can search by topic at http://lib.stat.cmu.edu/R/CRAN/

##Why packages?

Because they are useful.

Allow you to analyze particular data sets, or do specific (or generic types of processing)

Example: dataMaid

```{r}
#install.packages("dataMaid")
library(dataMaid)
library(ggplot2)
```
Note: make sure that you set your path to someplace useful (e.g., where you are storing notes for this class) in the last function. 

```{r}
makeDataReport(diamonds, file="/Users/farrahpl/Desktop/CS-347 Applied Data Science/week3/Diamonds.RMD") #diamonds is a dataset in ggplot2
```

## The insides of an R package 

Top level of an R package: folder named after the package 
   - common files:
      - DESCRIPTION: metadata about the project -- you don't create this by hand, RStudio/ devtools will create the template for you
      - NAMESPACE: defines the functions, classes, and methods that are imported into the name space. DO NOT WRITE THIS BY HAND
      - NEWS: where you put updates about the package; track changes between versions
      - CITATION: "Appropriate citations must me included in help pages and vignettes; this aspect of documentation is no different from any scientific endeavor" 
   - subfolders:
      - data -- stores your data: .rda, .csv, other formatted data
      - inst -- store other stuff that's not part of the "real" package. EX: JSON source files, other information about where/when you got the data 
      - R -- the folder that include your .R code 
      - man -- contains documentation for the package; do not write by hand Roxygen2 to auto-generate these 
      - vignettes -- where you put example markdown files for how to use your package
      - tests -- where unit tests for packages go 
      
##Creating R packages

First, install devtools:
```{r}
install.packages("devtools")
```

Then, create a new project:

RStudio -- File -- new project -- new directory -- R package with devtools

pick a directory, name your package ('PokemonPackage')

NOTE: R packages can only have names, numbers, and can be spaces but I recommend against this-- no underscores or other special characters 

RStudio will create the directory, and create the DESCRPTION and NAMESPACE files, as well as the R directory

Open the DESCRIPTION file; do some editing (e.g., your name, description, title ). Note that leading whitespace sometimes matters here, so you want to be careful with it.

Make a new folder in the package called 'data' go into the folder add the pokemon2.tsv file from 2u digital campus to the data directory.

Set the data dir as the working directory (it's an option under "more" in the lower right quadrant).

Run this code in the console:
```{r}
pokemon_data=read.csv("/Users/farrahpl/Desktop/CS-347 Applied Data Science/week3/pokemon2.tsv",sep="\t",stringsAsFactors = FALSE) #edit directory path
save(pokemon_data,file="pokemon_data.rda")
```

RDA files are the standard way that R stores data in a binary format. 
You can delete the tsv, or move it to the `inst` directory if you want to make one. 

Go back up to your package directory and set that back as your working directory.

Go to Build -- install and restart
You shouldn't get any errors, but you should see this:
```{r}
library(TestPokemonPackage)
```

In the console, type this:
```{r, echo=FALSE}
data(pokemon_data)
head(pokemon_data)
```
     
You can see the data!

## Creating functions 

Download PokemonPackage.R from the 2u digital campus, put in the R directory for your package. Open the file, and we'll start to fill in the functions.

Spend some time trying it out yourself!

Once you have one function done (or just want to test one that you're working on),
save the .R file, and build and restart the package again, then call the function from the console.

```{r}
#library(PokemonPackage)
#pokeDex("Bulbasaur")
```

Did you get an error message? Of course you did! A function has to be documented before it can be run.

##Documenting functions

### Why do we need documentation?

- Documentation is one of the most important aspects of a good package.
- Without it, users won’t know how to use your package.
- Documentation is also useful for future-you (so you remember what your functions were supposed to do), and for developers extending your package.

Writing R documentation by hand is a pain, so we use Roxygen for generating  documentation

How to use it:
   - add Roxygen comments (each line starts with #' )
   - Build the documentation
   - Roxygen will also do all the namespace stuff for you, which is what allows you to run your functions

Roxygen tags:
 @param name description
 @examples : Provide executable code to show how to use your function.
 @return description : Describes the output from the function.
 @importFrom:import selected functions, use this if you use any external packages (e.g. #'@importFrom mvtnorm pmvnorm where mvtnorm is the package and pmvnorm is the function we used) 
 @export : without this, people still can't use the function
  
 
Here is what we will put in our .R file:

```{r}
#' Function for getting data about a single pokemon
#' @param name name of the pokemon we want information about
#' @examples pokeDex("Charmander")
#' @examples pokeDex("Litwick")
#' @return None
#' @export
pokeDex<-function(name){
  print(head(pokemon_data)) #this will get changed later to do what the function is supposed to do

}
```

NOTE: spacing here is VERY IMPORTANT. You can't have any spaces before the \#s, there can only be one empty space after the ', and your function definition should come directly after your @export line.

Once you put the comments in, you need to use devtools or roxygen to actually build the documentation
  - one way to do it: Ctrl + shit + d
     (same as using Build --> document )
     
This will create your man folder in your package with your documentation for each of your functions. 
     
Once you build the documentation, build and install the package, then try to run the function again.

Also, now that we have documentation, we can do things like this:

```{r}
example(pokeDex)
?pokeDex
```

Way to go! You now have a package with data and functions!
      
